alternativeRSA

### 2016-01-20 models/RSA_alternativeCostFormulation.Rmd
added model run data with write to csv

Run RSA - five item scales
```{r}
computePosteriors <- c(norm.post, norm.post3)
# runFull() 
# ---------
# data       => contains scale | degree | stars | speaker.p | prior.p | cost
# calcFN     => without cost = 1, with cost = 2
# alpha      => set alpha level
runFull <- function(data, calcFn = 2, alpha = 1, numAlts = 5) {
  results <-
    data.frame(scale = c(),
               stars = c(),
               degree = c(),
               preds = c())

@@

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")

@@

# run fns()
checkSaliences()
tuneAlphas(modelData = data, humanDf = listener, numAlts = 2)
```

Data exploration Fns()
```{r}
# matchModelHumandData()
# -----------------------
# NEED TO FINISH WORKING HERE!!!!!!!!!!!!!!!!!!
matchModelHumanData <- function(predDf, listenerDf, study = "e12") {
  combinedDf <- data.frame(scale = c(),
                           degree = c(),
                           speaker.p = c(),
                           prior.p = c(),
                           cost = c(),
                           preds = c())

@@

Entailment analysis
```{r}

```

Negative alts analysis
```{r}

```

@@

  # plots
#   predPlot(combinedDf)
#   correlationScatterPlot(combinedDf)

@@

Entailment / Neg alts / Symmetric scale
```{r}
symmetricScalerun <- completeModelRun(speakerL12, listener, unifPriors, salienceCosts5, numAlts = 5)
symmetricScalerun[[4]]
negAltsRun <-        completeModelRun(speakerL10, listener, unifPriors, salienceCosts4, numAlts = 4)
negAltsRun[[4]]
entailmentRun <-      completeModelRun(speakerL08, listener, unifPriors, salienceCosts2, numAlts = 2)
entailmentRun[[4]]
```

Aggregate analysis
```{r}
cors <- data.frame(cors = c(symmetricScalerun[[4]], negAltsRun[[4]], entailmentRun[[4]]),
                   m = c("full", "neg_alts", "entailment"))
ggplot(cors, aes(x=reorder(m, cors), y = cors)) +
  geom_bar(stat="identity", position="dodge") +
  ylab("Correlation") +
  xlab("Model") +
  ylim(0, 1) + 
  ggtitle("Model fit correlation coefficients") +
  guides(colour=FALSE) +
  geom_text(aes(x=m, y=as.numeric(cors) + 0.025, label=round(cors, 3)))
```

```{r}
normedVec(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
#normedVec(subset(test, scale == "good_excellent" & degree == "mid")$preds)
sum(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
dfPreds <- test
dfHuman <- listener
name <- "good_excellent"
getCorrs <- function(dfPreds, dfHuman, name) {
  predSubset <- subset(dfPreds, scale == name & (degree == "hi1" | degree == "hi2"))
  predSubset$degree = with(predSubset,
                           ifelse(degree == "hi1", "hi", "low"))
  humanSubset <- subset(dfHuman, scale == name)
  joinedDf <- merge(x = predSubset,
                    y = humanSubset,
                    by = c("degree", "stars", "scale"),
                    all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))
  cor(joinedDf$preds, joinedDf$listener.p)
}

getCorrs(test, listener, "good_excellent")
getCorrs(test, listener, "liked_loved")
getCorrs(test, listener, "memorable_unforgettable")
getCorrs(test, listener, "palatable_delicious")

######################################################################
----------------------------------------------------------------------
######################################################################

runModel <- function(model, mat, degree, alpha) {
  if (model == 1) unlist(mapply(seq(1,5),
                                FUN=function(x) norm.post(x, degree, mat, alpha, priors)))
}
mat <- data %>% 
  select(stars, degree, speaker.p) %>%
  spread(degree, speaker.p) %>%
  mutate(hi1 = hi1 / sum(hi1),
         hi2 = hi2 / sum(hi2),
         mid = mid / sum(mid),
         low1 = low1 / sum(low1),
         low2 = low2 / sum(low2)) %>%
  select(hi1, hi2, mid, low1, low2)

runModel(1, mat[1:5, ], 1, 6)
runModel(1, mat[1:5, ], 2, 6)

subset(listener, scale == "good_excellent")

```

New cost formulation
```{r}
speaker.inf2 = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd2 = function(rating, degree, m, alpha, costs) {
  numerator = speaker.inf2(m[rating, degree], alpha, costs[degree])
  normalize = sum(unlist(sapply(seq(nrow(m)),
                                function(i) {speaker.inf2(m[i, ], alpha, costs[degree])})))
  return(numerator / normalize)
  # return(numerator)
}
# non-normalized posterior
# -----------------------
nn.post2 = function(rating, degree, m, alpha, costs, useprior) {
  prior = priors[rating, "prior.p"]
  # prior <- 0.2
  return(speaker.lhd2(rating, degree, m, alpha, costs) * prior)
}
# normalized posterior
# --------------------
norm.post2 = function(rating, degree, m, alpha, costs, useprior) {
  nn = nn.post2(rating, degree, m, alpha, costs, useprior)
  normalize = sum(unlist(sapply(seq(1, 5),
                                function(i){nn.post2(rating, i, m, alpha, costs, useprior)})))
  return(nn / normalize)
}
```

```{r}
# good excellent
mat[1:5,]
salience <- c(1, 1, 0.05, 0.08, 0.22)
normed.salience <- -1 * (salience / sum(salience))

normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[1:5,], 5, normed.salience))))
normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[1:5,], 5, normed.salience))))

#################################
# good_excellent
# --------------
# low
good.exc.low <-
  listener[with(listener, which(scale == "good_excellent" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 2))
# high
good.exc.high <-
  listener[with(listener, which(scale == "good_excellent" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 3))
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[1:5,], 7, normed.salience)))),
    good.exc.low[c(5, 4, 1, 2, 3)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[1:5,], 7, normed.salience)))),
    good.exc.high[c(5, 4, 3, 1, 2)])
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[1:5,], 7)))),
    good.exc.low[c(5, 4, 1, 2, 3)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[1:5,], 7)))),
    good.exc.high[c(5, 4, 3, 1, 2)])

#################################
# liked_loved
# low
liked.loved.low <-
  listener[with(listener, which(scale == "liked_loved" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 2))
# high
liked.loved.high <-
  listener[with(listener, which(scale == "liked_loved" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 3))
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[6:10,], 7, normed.salience)))),
    liked.loved.low[c(5, 1, 2, 3, 4)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[6:10,], 7, normed.salience)))),
    liked.loved.high[c(5, 4, 3, 1, 2)])
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[6:10,], 7)))),
    liked.loved.low[c(5, 1, 2, 3, 4)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[6:10,], 7)))),
    liked.loved.high[c(5, 4, 3, 1, 2)])

#################################
# memorable_unforgettable
# low
mem.unf.low <-
  listener[with(listener, which(scale == "memorable_unforgettable" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 3))
mem.unf.low <- mem.unf.low[c(5, 4, 3, 2, 1)]
# high
mem.unf.high <-
  listener[with(listener,
                which(scale == "memorable_unforgettable" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 1))
mem.unf.high <- mem.unf.high[c(1, 5, 2, 3, 4)]
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[11:15,], 7, normed.salience)))),
    mem.unf.low)
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[6:10,], 7, normed.salience)))),
    mem.unf.high)
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[6:10,], 3)))),
    mem.unf.low)
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[6:10,], 7)))),
    mem.unf.high)
```

How to implement this using saliency rewards (negative costs)???
```{r}
run.newCosts = function(d, alpha=1, useprior=F, usenone=F, normalize=F) {
  mat <- data.full.extras %>% 
    select(stars, degree, speaker.p) %>%
    spread(degree, speaker.p) %>%
    select(hi1, hi2, mid, low1, low2)

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  } 
  d$degree <- with(d, ifelse(degree == "hi1", 5,
                             ifelse(degree == "hi2", 4,
                                    ifelse(degree == "mid", 3, 
                                           ifelse(degree == "low1", 2, 3)))))
  d$pred = round(as.numeric(mapply(norm.post2, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha,
                                                    costs = normed.salience,
                                                    useprior = useprior))), 
                  digits=4)

  return(normalize(d))
}
run.newCosts(data.full.extras, alpha=7, useprior=TRUE, usenone=FALSE, normalize=normalize)
```

Old `run` function
```{r}
run.full = function(d, alpha=1, useprior=F, usenone=F, addMid=F, normalize=F) {
  # alpha = scales.entropy[scales.entropy$scale==d$scale[1], ]$Entropy
  if (addMid) {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               mid = mid / sum(mid),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, mid, low1, low2)    
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, mid, low1, low2)    
    }
  } else {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, low1, low2)
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, low1, low2)
    }
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}
```

```{r}
run.full(data.full.extras, useprior = TRUE, usenone = FALSE, addMid = TRUE, normalize = TRUE)
cor(data.full.extras$listener.p, data.full.extras$pred)
```

>>>>

Run RSA
```{r}
computePosteriors <- c(norm.post, norm.post3)
# runFull() 
# ---------
# data       => contains scale | degree | stars | speaker.p | prior.p | cost
# calcFN     => without cost = 1, with cost = 2
# alpha      => set alpha level
runFull <- function(data, calcFn = 2, alpha = 1, numAlts = 5) {
  results <-
    data.frame(scale = c(),
               stars = c(),
               degree = c(),
               preds = c())

@@

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)

@@

# run fns()
# ---------
# checkSaliences()
# tuneAlphas(modelData = data, humanDf = listener, numAlts = 2)
```

Data exploration Fns()
```{r}
# matchModelHumandData()
# -----------------------
matchModelHumanData <- function(predDf, listenerDf, study = "e12") {
  combinedDf <- data.frame(scale = c(),
                           degree = c(),
                           speaker.p = c(),
                           prior.p = c(),
                           cost = c(),
                           preds = c())

@@

@@

@@

Entailment / Neg alts / Symmetric scale
```{r}
symmetricScalerun <- completeModelRun(speakerL12, listener, unifPriors, salienceCosts5, numAlts = 5)
symmetricScalerun[[4]]
negAltsRun <- completeModelRun(speakerL10, listener, unifPriors, salienceCosts4, numAlts = 4)
negAltsRun[[4]]
entailmentRun <- completeModelRun(speakerL08, listener, unifPriors, salienceCosts2, numAlts = 2)
entailmentRun[[4]]

######WORKING HERE!!!!!!

listener2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv", stringsAsFactors = FALSE)
symmetricData <- symmetricScalerun[[1]]
symmetricData$numAlts = 5
negAltsData <- negAltsRun[[1]]
negAltsData$numAlts = 4
entailmentData <- entailmentRun[[1]]
entailmentData$numAlts = 2

# Aggregate data to write to .csv

# combine with e11 prag listener nums
aggregateData <- rbind(symmetricData, negAltsData, entailmentData)
listener2 <- listener2[(listener2$scale == "some_all" & (listener2$degree == "low1" | listener2$degree == "hi1")) |
                         (listener2$scale != "some_all" & (listener2$degree == "hi1" | listener2$degree == "hi2")), ]
listener2$degree <- as.character(ifelse(listener2$degree == "hi1", "hi", "low"))
names(listener2) <- c("degree", "scale", "stars", "listener.p2", "total")
aggregateData <- left_join(aggregateData, listener2[, -5], by=c("degree", "stars", "scale")) %>%
  mutate(listener.p2 = ifelse(is.na(listener.p2), 0, listener.p2))

# record correlations for both studies
modelRunData <- aggregateData
symmetricE6cor <- cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p)
symmetricE11cor <-cor(subset(modelRunData, numAlts == 5)$preds, subset(modelRunData, numAlts == 5)$listener.p2)
negAltsE6cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p)
negAltsE11cor <- cor(subset(modelRunData, numAlts == 4)$preds, subset(modelRunData, numAlts == 4)$listener.p2)
entailmentE6cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p)
entailmentE11cor <- cor(subset(modelRunData, numAlts == 2)$preds, subset(modelRunData, numAlts == 2)$listener.p2)
modelRunData$corE6 <- ifelse(modelRunData$numAlts == 5, symmetricE6cor,
                             ifelse(modelRunData$numAlts == 4, negAltsE6cor, entailmentE6cor))
modelRunData$corE11 <- ifelse(modelRunData$numAlts == 5, symmetricE11cor,
                             ifelse(modelRunData$numAlts == 4, negAltsE11cor, entailmentE11cor))
modelRunData$alphas <- ifelse(modelRunData$numAlts == 5, symmetricScalerun[[3]],
                             ifelse(modelRunData$numAlts == 4, negAltsRun[[3]], entailmentRun[[3]]))

# write.csv(modelRunData, "model_data/allModelRuns.csv")
```
